---
title: "r-ambidexterity"
author: "Charreau Bell"
output: html_notebook
---

# Translating Between R and Python
Let's look at an example of some of the ways that we can leverage the underlying principles of R to make better use of tidyverse functions.

```{r relevant libraries}
library(tidyverse)
```


# Data Analysis
We have the following data summarizing the sports performances of a few players.  We want to perform some analysis to understand how the players are playing.

## Jump right in
```{r generate data}
#we'll use some custom-created data for simplicity
sports_df <- tibble(player_name = c('Johnson', 'James', 'Aniston', 'Errol', 'Caine', 'Jamison'),
                     team = c('Tigers', 'Tigers', 'Lions', 'Lions', 'Lions', 'Bears'),
                     game_1 = c(12.2, 3.3, 32.5, 10.4, 8.45, 5.45),
                     game_2 = c(34, 45, 23, 13, 8, 24),
                     game_3 = c(14, 23, 76, 19, 26, 53))

#let's view it
sports_df
```

### Calculate the mean points per game
```{r mean points per game}

```

### Calculate column with mean points per player
```{r mean points per player}

```

### Plot the overall game point distributions
```{r plot overall game points distribution}

```

**What are the challenges of this approach?**

**How does the data and approach diverge from tidyverse principles?**

**What is an alternate approach which is more consistent with the tidyverse principles?**

## Alternate approach
```{r tidy the data}

```

```{r tidy mean points per game}

```
```{r tidy mean points per player}

```

```{r tidy plot overall game points distribution}

```

**What were the differences between this and the first approach?  What indicators were there that this approach aligns with tidyverse principles?**






